On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-05-17 Manuscript/manuscript.tex
cleaning/comments

\begin{abstract}
Block bootstrap is widely used in constructing confidence intervals for
parameters estimated from stationary time series. Theoretically, the method
should provide valid confidence intervals as the length of the time series goes
to infinity. In practice, however, it is necessary to know how large of a finite
sample is required for block bootstrap confidence intervals to work well. This
study aims to answer this question in a simple simulation setting where the data
are generated from an autoregressive process of order~1. The empirical coverage
rates of several commonly used bootstrap confidence intervals for the mean,
standard deviation, and the lag-1 autocorrelation are compared. \jy{put most
  notable findings here. Possibly 1) large sample is needed; percentile
  bootstrap is not working well for rho (can we offer some explanation?).}

@@

@@

\subsection{Block Bootstrap Sample}

@@

Alternatively, block-bootstrap can be done with overlapping or moving blocks.
Define moving blocks $Z_j = \{X_j, \ldots, X_{j + l - 1}\}$,
$j = 1, \ldots, n - l + 1$. Now we draw $k$ blocks from the $(n - l + 1)$
blocks
of $Z_j$'s with replacement and then align them in the order they were
picked
to form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last
block selected will be reduced in size so that the final size of the
block bootstrap sample is $n$. It is also possible to implement moving block
bootstrap while allowing blocks to wrap around the end of the series. In other
words, define moving blocks (assuming $l > 1$) as:
\begin{equation}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & \text{if } j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & \text{if } j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation}
The wrap-around version does not require that $n/l$ be an integer.
We used this version in this study as implemented in the function
\texttt{tsboot} from R package \textsl{boot} \citep{boot}.

@@

\jy{order them as we discussed.}
\mc{addressed}
\paragraph{Standard Normal CI}
The procedure for constructing a Standard Normal CI is described in 
\citet{efron1993introduction}.
A $(1 - \alpha$ standard normal CI depends on the standard error $\hat{SE}$
for each of the
$B$ bootstrap point estimates
$\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$. 
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\hat{SE}, \hat{\theta}_{n} -
z_{(\alpha/2)}\hat{SE}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
The standard
normal CI is a table-based bootstrap interval approach, which is to say that
it relies on the 
assumption that $\hat\theta_n$ is normally distributed and the expected value
of $\hat\theta_n$
is $\theta$.

@@

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
\eds{a different abbrevation/notation was used for this in the introduction}
\mc{used $BC_a$}
\jy{Isn't this just the empirical CDF? We just need the sample quantile functon.
  What is your reference for this description?}
Let $\hat{z}_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$
for $b \in \{1, \ldots, B\}$. 
Let $Z_{(i)}$ be the original sample without the $i$th block $z_i$, and let
$\hat{\theta}_{n, -i}$ be the jackknife values of $Z_{(i)}$
for $i \in \{1, \ldots, n\}$, 
and let $a$ be $SKEW\{\hat{\theta}_{n, -i}\} / 6$. Then
$\hat\theta_{BC_a}[\alpha] = \hat{G}^{-1}\Phi(z_0 + \frac{z_{0} +
  z_{\alpha}}{1 - a(z_{0} + z_{\alpha})}).$
\eds{notation will need to be cleaned and clarified}
A $(1 - \alpha)$ centered $BC_a$ CI is
\[
(\hat\theta_{BC_a}[\alpha/2], \quad
\hat\theta_{BC_a}[1 - \alpha/2]).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. The acceleration factor refers to
the 
rate of change of of the standard error of $\hat\theta_n$ with respect to
$\theta$.
\jy{why is it called BCA?}
\mc{addressed}

@@

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x = 1$, the standard deviation of $X_t$; and
2) $\phi$, the lag-1 auto-correlation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, 400, 800, 1600, 3200\}$. In each configuration, we generated 10000
replicates. For each replicate, we constructed three 95\% block bootstrap
CIs for each parameter as described in the last section. We can estimate 
$\mu$, $\sigma_x$, and $\phi$ by computing the sample mean, sample 
standard deviation,
and lag-1 auto-correlation, respectively, of each bootstrap sample, and then 
constructing intervals for each parameter using the appropriate procedures
described in Section~\ref{sec:bbci} .
Then we estimated their actual coverage rates along with their 95\% confidence
intervals from the 10000 replicates. The block bootstrap sampling step was done
with function \texttt{tsboot} from R package \textsl{boot} \citep{boot}, with
block size $\lceil n / l \rceil$. This function by default is an implementation
of moving block bootstrap as described in the previous section, meaning that
that blocks are allowed to wrap around, and $l = \lceil n^{1/3} \rceil$.
\jy{size? also it was overlapping blocks. Describe the details here.}
\mc{addressed}
\eds{Is $l = \lceil n^{1/3} \rceil$? Is it also worth investigating different 
choices of $l$ for fixed (large) $n$?}
\mc{Addressed in the Block Bootstrap Sample subsection}
\eds{how did you estimate $\mu$, $\sigma_x$ and $\phi$?}
\mc{addressed}

@@

\jy{One paragraph for each parameter; write down your observations in bullet
  points and then organize them into a logically flowing paragraph.}
\mc{paragraphs below}

@@

\jy{One summary paragraph at the end.}
\mc{paragraph below}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_mu}
  \caption{Plots of $\mu$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600, 3200$ were recorded.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_sigma}
  \caption{Plots of $\sigma_x$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600, 3200$ were recorded.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_phi}
  \caption{Plots of $\phi$ coverage rates for each interval type and $\phi$
    combination. The orange dotted line represents the 95\% nominal coverage
    rate. The error bars represent a 95\% CI for the coverage
    proportion out of 10000 replications for each $n$. Results for 
    $n = 100, 200, 400, 800, 1600, 3200$ were recorded.}
  \label{fig:phi}
\end{figure}

>>>>

\begin{abstract}
Block bootstrap is widely used in constructing confidence intervals for
parameters estimated from stationary time series. Theoretically, the method
should provide valid confidence intervals as the length of the time series goes
to infinity. In practice, however, it is necessary to know how large of a finite
sample is required for block bootstrap confidence intervals to work well. This
study aims to answer this question in a simple simulation setting where the data
are generated from a first-order autoregressive process. The empirical coverage
rates of several commonly used bootstrap confidence intervals for the mean,
standard deviation, and the lag-1 autocorrelation are compared. \jy{put most
  notable findings here. Possibly 1) large sample is needed; percentile
  bootstrap is not working well for rho (can we offer some explanation?).}

@@

\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

@@

\subsection{Block Bootstrap}

@@

Alternatively, block-bootstrap can be done with overlapping or moving blocks.
Define moving blocks $Z_j = \{X_j, \ldots, X_{j + l - 1}\}$,
$j = 1, \ldots, n - l + 1$. Now we draw $k$ blocks from the $(n - l + 1)$
blocks
of $Z_j$'s with replacement and then align them in the order they were
picked
to form a block bootstrap sample. If $n$ is not a multiple of~$l$, the last
block selected will be reduced in size so that the final size of the
block bootstrap sample is $n$. It is also possible to implement moving block
bootstrap while allowing blocks to wrap around the end of the series. In other
words, define moving blocks (assuming $l > 1$) as:
\begin{equation}
Z_j =
    \begin{cases}
        \{X_j, \ldots, X_{j + l - 1}\}, & \text{if } j = 1, \dots, n - l + 1,\\
        \{X_j, \ldots, X_n, X_1, \ldots, X_{j-n+l-1}\}, & \text{if } j = n - l
        + 2 ,\dots, n.
    \end{cases}
\end{equation}
This version does not require that $n/l$ be an integer.
We used this version in this study as implemented in the function
\texttt{tsboot} from R package \textsl{boot} \citep{boot}.

@@

\jy{Read carpenter2000bootstrap and cite it; quite useful in sorting the literature}

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$.
A $(1 - \alpha)$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonaly
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

@@

\paragraph{Bias Corrected and Accelerated ($BC_a$) Bootstrap CI}
The $BC_a$ CI was first suggested in \citet{efron1987better}.
\eds{a different abbrevation/notation was used for this in the introduction}
\mc{used $BC_a$}
\jy{Isn't this just the empirical CDF? We just need the sample quantile functon.
  What is your reference for this description?}
Let $\hat{z}_0 = \Phi^{-1}\{\#\{\hat\theta_n^{(b)} < \hat{\theta}_n\} / B\}$
for $b \in \{1, \ldots, B\}$. 
Let $Z_{(i)}$ be the original sample without the $i$th block $z_i$, and let
$\hat{\theta}_{n, -i}$ be the jackknife values of $Z_{(i)}$
for $i \in \{1, \ldots, n\}$, 
and let $a$ be $SKEW\{\hat{\theta}_{n, -i}\} / 6$. Then
$\hat\theta_{BC_a}[\alpha] = \hat{G}^{-1}\Phi(z_0 + \frac{z_{0} +
  z_{\alpha}}{1 - a(z_{0} + z_{\alpha})}).$
\eds{notation will need to be cleaned and clarified}
A $(1 - \alpha)$ centered $BC_a$ CI is
\[
(\hat\theta_{BC_a}[\alpha/2], \quad
\hat\theta_{BC_a}[1 - \alpha/2]).
\]
This CI is not necessarily centered by the point estimator $\hat\theta_n$. The
$BC_a$ method corrects for bias and skewness of the $B$ bootstrap point
estimates $\{\hat\theta_n^{(1)}, \ldots, \hat\theta_n^{(B)}\}$ by including
bias-correction and acceleration factors. The acceleration factor refers to
the 
rate of change of of the standard error of $\hat\theta_n$ with respect to
$\theta$.

@@

Three target parameters of $X_t$ were considered:
1) $\mu = 0$, the mean of $X_t$;
2) $\sigma_x = 1$, the standard deviation of $X_t$; and
2) $\phi$, the lag-1 auto-correlation coefficient.
To investigate the effect of sample size~$n$, we considered an array of values
$n \in \{100, 200, 400, 800, 1600, 3200\}$. In each configuration, we generated 10000
replicates. For each replicate, we constructed three 95\% block bootstrap
CIs for each parameter as described in the last section. We can estimate 
$\mu$, $\sigma_x$, and $\phi$ by computing the sample mean, sample 
standard deviation,
and lag-1 auto-correlation, respectively, of each bootstrap sample, and then 
constructing intervals for each parameter using the appropriate procedures
described in Section~\ref{sec:bbci} .
Then we estimated their actual coverage rates along with their 95\% confidence
intervals from the 10000 replicates. The block bootstrap sampling step was done
with function \texttt{tsboot} from R package \textsl{boot} \citep{boot}, with
block size $\lceil n / l \rceil$. This function by default is an implementation
of moving block bootstrap as described in the previous section, meaning that
that blocks are allowed to wrap around, and $l = \lceil n^{1/3} \rceil$.

\eds{Is $l = \lceil n^{1/3} \rceil$? Is it also worth investigating different 
choices of $l$ for fixed (large) $n$?}
\mc{Addressed in the Block Bootstrap Sample subsection}
\eds{how did you estimate $\mu$, $\sigma_x$ and $\phi$?}
\mc{addressed}

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_mu}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates.
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:mu}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_sigma}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma$ of an AR(1) process with AR
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates.
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_phi}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates.
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi}
\end{figure}

\jy{I'm curiouss why the coverge is so poor for the percentile method. Could
  plot 50 intervals for phi using different sample sizes?}

@@

@@
